%% grace-3.tex - Part 2c: Of Man's Redemption (Christ's Suffering and Death)
%% Questions 37-45, Lord's Days XV-XVII

%% ============================================================
%% LORD'S DAY XV
%% ============================================================

\lordsday{15}

\catechismqa{37}{%
What do you confess when you say that he suffered?%
}{%
During all the time he lived on earth, but especially at the end, Christ bore in body and soul the wrath of God against the sin of the whole human race. Thus, by his suffering, as the only atoning sacrifice, he has redeemed our body and soul from everlasting damnation, and obtained for us the grace of God, righteousness, and eternal life.%
}

\catechismqa{38}{%
Why did he suffer under Pontius Pilate as judge?%
}{%
Though innocent, Christ was condemned by an earthly judge, and so he freed us from the severe judgment of God that was to fall on us.%
}

\catechismqa{39}{%
Does it have a special meaning that Christ was crucified and did not die in a different way?%
}{%
Yes. Thereby I am assured that he took upon himself the curse which lay on me, for a crucified one was cursed by God.%
}

%% ============================================================
%% LORD'S DAY XVI
%% ============================================================

\vspace{0.5em}
\lordsday{16}

\catechismqa{40}{%
Why was it necessary for Christ to humble himself even unto death?%
}{%
Because of the justice and truth of God satisfaction for our sins could be made in no other way than by the death of the Son of God.%
}

\catechismqa{41}{%
Why was he buried?%
}{%
His burial testified that he had really died.%
}

\catechismqa{42}{%
Since Christ has died for us, why do we still have to die?%
}{%
Our death is not a payment for our sins, but it puts an end to sin and is an entrance into eternal life.%
}

\catechismqa{43}{%
What further benefit do we receive from Christ's sacrifice and death on the cross?%
}{%
Through Christ's death our old nature is crucified, put to death, and buried with him, so that the evil desires of the flesh may no longer reign in us, but that we may offer ourselves to him as a sacrifice of thankfulness.%
}

\catechismqa{44}{%
Why is there added: He descended into hell?%
}{%
In my greatest sorrows and temptations I may be assured and comforted that my Lord Jesus Christ, by his unspeakable anguish, pain, terror, and agony, which he endured throughout all his sufferings but especially on the cross, has delivered me from the anguish and torment of hell.%
}

%% ============================================================
%% LORD'S DAY XVII
%% ============================================================

\vspace{0.5em}
\lordsday{17}

\catechismqa{45}{%
How does Christ's resurrection benefit us?%
}{%
First, by his resurrection he has overcome death, so that he could make us share in the righteousness which he had obtained for us by his death. Second, by his power we too are raised up to a new life. Third, Christ's resurrection is to us a sure pledge of our glorious resurrection.%
}
