%% grace-part-c.tex - Part 2c: Of Man's Redemption
%% Questions 50-68, Lord's Days XIX-XXV (19 questions)

%% ============================================================
%% LORD'S DAY XIX
%% ============================================================

\lordsday{19}

\catechismqa{50}{%
Why is it added, and sits at the right hand of God?%
}{%
Christ ascended into heaven to manifest himself there as Head of his church, through whom the Father governs all things.%
}

\catechismqa{51}{%
How does the glory of Christ, our Head, benefit us?%
}{%
First, by his Holy Spirit he pours out heavenly gifts upon us, his members. Second, by his power he defends and preserves us against all enemies.%
}

\catechismqa{52}{%
What comfort is it to you that Christ will come to judge the living and the dead?%
}{%
In all my sorrow and persecution I lift up my head and eagerly await as judge from heaven the very same person who before has submitted himself to the judgment of God for my sake, and has removed all the curse from me. He will cast all his and my enemies into everlasting condemnation, but he will take me and all his chosen ones to himself into heavenly joy and glory.%
}

%% ============================================================
%% LORD'S DAY XX
%% ============================================================

\vspace{0.5em}
\lordsday{20}

\catechismqa{53}{%
What do you believe concerning the Holy Spirit?%
}{%
First, he is, together with the Father and the Son, true and eternal God. Second, he is also given to me, to make me by true faith share in Christ and all his benefits, to comfort me, and to remain with me forever.%
}

%% ============================================================
%% LORD'S DAY XXI
%% ============================================================

\vspace{0.5em}
\lordsday{21}

\catechismqa{54}{%
What do you believe concerning the holy catholic church?%
}{%
I believe that the Son of God, out of the whole human race, from the beginning of the world to its end, gathers, defends, and preserves for himself, by his Spirit and Word, in the unity of the true faith, a church chosen to everlasting life. And I believe that I am and forever shall remain a living member of it.%
}

\catechismqa{55}{%
What do you understand by the communion of saints?%
}{%
First, that believers, all and everyone, as members of Christ have communion with him and share in all his treasures and gifts. Second, that everyone is bound to use his gifts readily and cheerfully for the benefit and well-being of the other members.%
}

\catechismqa{56}{%
What do you believe concerning the forgiveness of sins?%
}{%
I believe that God, because of Christ's satisfaction, will no longer remember my sins, nor my sinful nature, against which I have to struggle all my life, but will graciously grant me the righteousness of Christ, that I may never come into condemnation.%
}

%% ============================================================
%% LORD'S DAY XXII
%% ============================================================

\vspace{0.5em}
\lordsday{22}

\catechismqa{57}{%
What comfort does the resurrection of the body offer you?%
}{%
Not only shall my soul after this life immediately be taken up to Christ, my Head, but also this my flesh, raised by the power of Christ, shall be reunited with my soul and made like the glorious body of Christ.%
}

\catechismqa{58}{%
What comfort do you receive from the article about life everlasting?%
}{%
Even as I now already feel in my heart the beginning of eternal joy, so after this life I shall have perfect blessedness, such as no eye has seen, nor ear heard, nor the heart of man conceived---a blessedness in which to praise God forever.%
}

%% ============================================================
%% LORD'S DAY XXIII
%% ============================================================

\vspace{0.5em}
\lordsday{23}

\catechismqa{59}{%
But what does it help you now that you believe all this?%
}{%
In Christ I am righteous before God and an heir to life everlasting.%
}

\catechismqa{60}{%
How are you righteous before God?%
}{%
Only by true faith in Jesus Christ. Although my conscience accuses me that I have grievously sinned against all God's commandments, have never kept any of them, and am still inclined to all evil, yet God, without any merit of my own, out of mere grace, grants and credits to me the perfect satisfaction, righteousness, and holiness of Christ, as if I had never sinned nor been a sinner, and as if I had been as perfectly obedient as Christ was obedient for me---if only I accept this gift with a believing heart.%
}

\catechismqa{61}{%
Why do you say that you are righteous only by faith?%
}{%
Not that I am acceptable to God on account of the worthiness of my faith, for only the satisfaction, righteousness, and holiness of Christ is my righteousness before God. I can receive this righteousness and make it my own by faith only.%
}

%% ============================================================
%% LORD'S DAY XXIV
%% ============================================================

\vspace{0.5em}
\lordsday{24}

\catechismqa{62}{%
But why can our good works not be our righteousness before God, or at least a part of it?%
}{%
Because the righteousness which can stand before God's judgment must be absolutely perfect and wholly in agreement with the law of God, whereas even our best works in this life are all imperfect and defiled with sin.%
}

\catechismqa{63}{%
But do our good works earn nothing, even though God promises to reward them in this life and the next?%
}{%
This reward is not earned; it is a gift of grace.%
}

\catechismqa{64}{%
Does this teaching not make people careless and wicked?%
}{%
No. It is impossible that those who are grafted into Christ by true faith should not bring forth fruits of thankfulness.%
}

%% ============================================================
%% LORD'S DAY XXV
%% ============================================================

\vspace{0.5em}
\lordsday{25}

\catechismqa{65}{%
Since then faith alone makes us share in Christ and all his benefits, where does this faith come from?%
}{%
From the Holy Spirit, who works it in our hearts by the preaching of the gospel, and strengthens it by the use of the sacraments.%
}

\catechismqa{66}{%
What are the sacraments?%
}{%
The sacraments are visible, holy signs and seals instituted by God so that by their use he may the more fully declare and seal to us the promise of the gospel: that because of the one sacrifice of Christ accomplished on the cross he grants us freely the forgiveness of all our sins and eternal life.%
}

\catechismqa{67}{%
Are both the Word and the sacraments intended to focus our faith on the sacrifice of Jesus Christ on the cross as the only ground of our salvation?%
}{%
Yes, indeed. The Holy Spirit teaches us in the gospel and assures us by the sacraments that our entire salvation rests on the one sacrifice of Christ made for us on the cross.%
}

\catechismqa{68}{%
How many sacraments has Christ instituted in the new covenant?%
}{%
Two: holy baptism and the holy supper.%
}
