%% grace-part-b.tex - Part 2b: Of Man's Redemption
%% Questions 31-49, Lord's Days XII-XVIII (19 questions)

%% ============================================================
%% LORD'S DAY XII
%% ============================================================

\lordsday{12}

\catechismqa{31}{%
Why is he called Christ, that is, Anointed?%
}{%
Because he has been ordained by God the Father, and anointed with the Holy Spirit, to be our chief Prophet and Teacher, who has fully revealed to us the secret counsel and will of God concerning our redemption; our only High Priest, who by the one sacrifice of his body has redeemed us, and who continually intercedes for us before the Father; and our eternal King, who governs us by his Word and Spirit, and who defends and preserves us in the redemption obtained for us.%
}

\catechismqa{32}{%
Why are you called a Christian?%
}{%
Because I am a member of Christ by faith and thus share in his anointing, so that I may as prophet confess his name, as priest present myself a living sacrifice of thankfulness to him, and as king fight with a free and good conscience against sin and the devil in this life, and hereafter reign with him eternally over all creatures.%
}

%% ============================================================
%% LORD'S DAY XIII
%% ============================================================

\vspace{0.5em}
\lordsday{13}

\catechismqa{33}{%
Why is he called God's only-begotten Son, since we also are children of God?%
}{%
Because Christ alone is the eternal, natural Son of God. We, however, are children of God by adoption, through grace, for Christ's sake.%
}

\catechismqa{34}{%
Why do you call him our Lord?%
}{%
Because he has ransomed us, body and soul, from all our sins, not with silver or gold but with his precious blood, and has freed us from all the power of the devil to make us his own possession.%
}

%% ============================================================
%% LORD'S DAY XIV
%% ============================================================

\vspace{0.5em}
\lordsday{14}

\catechismqa{35}{%
What do you confess when you say: He was conceived by the Holy Spirit, born of the virgin Mary?%
}{%
The eternal Son of God, who is and remains true and eternal God, took upon himself true human nature from the flesh and blood of the virgin Mary, through the working of the Holy Spirit. Thus he is also the true seed of David, and like his brothers in every respect, yet without sin.%
}

\catechismqa{36}{%
What benefit do you receive from the holy conception and birth of Christ?%
}{%
He is our Mediator, and with his innocence and perfect holiness covers, in the sight of God, my sin, in which I was conceived and born.%
}

%% ============================================================
%% LORD'S DAY XV
%% ============================================================

\vspace{0.5em}
\lordsday{15}

\catechismqa{37}{%
What do you confess when you say that he suffered?%
}{%
During all the time he lived on earth, but especially at the end, Christ bore in body and soul the wrath of God against the sin of the whole human race. Thus, by his suffering, as the only atoning sacrifice, he has redeemed our body and soul from everlasting damnation, and obtained for us the grace of God, righteousness, and eternal life.%
}

\catechismqa{38}{%
Why did he suffer under Pontius Pilate as judge?%
}{%
Though innocent, Christ was condemned by an earthly judge, and so he freed us from the severe judgment of God that was to fall on us.%
}

\catechismqa{39}{%
Does it have a special meaning that Christ was crucified and did not die in a different way?%
}{%
Yes. Thereby I am assured that he took upon himself the curse which lay on me, for a crucified one was cursed by God.%
}

%% ============================================================
%% LORD'S DAY XVI
%% ============================================================

\vspace{0.5em}
\lordsday{16}

\catechismqa{40}{%
Why was it necessary for Christ to humble himself even unto death?%
}{%
Because of the justice and truth of God satisfaction for our sins could be made in no other way than by the death of the Son of God.%
}

\catechismqa{41}{%
Why was he buried?%
}{%
His burial testified that he had really died.%
}

\catechismqa{42}{%
Since Christ has died for us, why do we still have to die?%
}{%
Our death is not a payment for our sins, but it puts an end to sin and is an entrance into eternal life.%
}

\catechismqa{43}{%
What further benefit do we receive from Christ's sacrifice and death on the cross?%
}{%
Through Christ's death our old nature is crucified, put to death, and buried with him, so that the evil desires of the flesh may no longer reign in us, but that we may offer ourselves to him as a sacrifice of thankfulness.%
}

\catechismqa{44}{%
Why is there added: He descended into hell?%
}{%
In my greatest sorrows and temptations I may be assured and comforted that my Lord Jesus Christ, by his unspeakable anguish, pain, terror, and agony, which he endured throughout all his sufferings but especially on the cross, has delivered me from the anguish and torment of hell.%
}

%% ============================================================
%% LORD'S DAY XVII
%% ============================================================

\vspace{0.5em}
\lordsday{17}

\catechismqa{45}{%
How does Christ's resurrection benefit us?%
}{%
First, by his resurrection he has overcome death, so that he could make us share in the righteousness which he had obtained for us by his death. Second, by his power we too are raised up to a new life. Third, Christ's resurrection is to us a sure pledge of our glorious resurrection.%
}

%% ============================================================
%% LORD'S DAY XVIII
%% ============================================================

\vspace{0.5em}
\lordsday{18}

\catechismqa{46}{%
What do you confess when you say, he ascended into heaven?%
}{%
That Christ, before the eyes of his disciples, was taken up from the earth into heaven, and that he is there for our benefit until he comes again to judge the living and the dead.%
}

\catechismqa{47}{%
Is Christ, then, not with us until the end of the world, as he has promised us?%
}{%
Christ is true man and true God. With respect to his human nature he is no longer on earth, but with respect to his divinity, majesty, grace, and Spirit he is never absent from us.%
}

\catechismqa{48}{%
But are the two natures in Christ not separated from each other if his human nature is not present wherever his divinity is?%
}{%
Not at all, for his divinity has no limits and is present everywhere. So it must follow that his divinity is indeed beyond the human nature which he has taken on and nevertheless is within this human nature and remains personally united with it.%
}

\catechismqa{49}{%
How does Christ's ascension into heaven benefit us?%
}{%
First, he is our Advocate in heaven before his Father. Second, we have our flesh in heaven as a sure pledge that he, our Head, will also take us, his members, up to himself. Third, he sends us his Spirit as a counter-pledge, by whose power we seek the things that are above, where Christ is, seated at the right hand of God, and not the things that are on earth.%
}
