%% grace-4.tex - Part 2d: Of Man's Redemption (Christ's Ascension and Glory)
%% Questions 46-56, Lord's Days XVIII-XXI

%% ============================================================
%% LORD'S DAY XVIII
%% ============================================================

\lordsday{18}

\catechismqa{46}{%
What do you confess when you say, he ascended into heaven?%
}{%
That Christ, before the eyes of his disciples, was taken up from the earth into heaven, and that he is there for our benefit until he comes again to judge the living and the dead.%
}

\catechismqa{47}{%
Is Christ, then, not with us until the end of the world, as he has promised us?%
}{%
Christ is true man and true God. With respect to his human nature he is no longer on earth, but with respect to his divinity, majesty, grace, and Spirit he is never absent from us.%
}

\catechismqa{48}{%
But are the two natures in Christ not separated from each other if his human nature is not present wherever his divinity is?%
}{%
Not at all, for his divinity has no limits and is present everywhere. So it must follow that his divinity is indeed beyond the human nature which he has taken on and nevertheless is within this human nature and remains personally united with it.%
}

\catechismqa{49}{%
How does Christ's ascension into heaven benefit us?%
}{%
First, he is our Advocate in heaven before his Father. Second, we have our flesh in heaven as a sure pledge that he, our Head, will also take us, his members, up to himself. Third, he sends us his Spirit as a counter-pledge, by whose power we seek the things that are above, where Christ is, seated at the right hand of God, and not the things that are on earth.%
}

%% ============================================================
%% LORD'S DAY XIX
%% ============================================================

\vspace{0.5em}
\lordsday{19}

\catechismqa{50}{%
Why is it added, and sits at the right hand of God?%
}{%
Christ ascended into heaven to manifest himself there as Head of his church, through whom the Father governs all things.%
}

\catechismqa{51}{%
How does the glory of Christ, our Head, benefit us?%
}{%
First, by his Holy Spirit he pours out heavenly gifts upon us, his members. Second, by his power he defends and preserves us against all enemies.%
}

\catechismqa{52}{%
What comfort is it to you that Christ will come to judge the living and the dead?%
}{%
In all my sorrow and persecution I lift up my head and eagerly await as judge from heaven the very same person who before has submitted himself to the judgment of God for my sake, and has removed all the curse from me. He will cast all his and my enemies into everlasting condemnation, but he will take me and all his chosen ones to himself into heavenly joy and glory.%
}

%% ============================================================
%% LORD'S DAY XX
%% ============================================================

\vspace{0.5em}
\lordsday{20}

\catechismqa{53}{%
What do you believe concerning the Holy Spirit?%
}{%
First, he is, together with the Father and the Son, true and eternal God. Second, he is also given to me, to make me by true faith share in Christ and all his benefits, to comfort me, and to remain with me forever.%
}

%% ============================================================
%% LORD'S DAY XXI
%% ============================================================

\vspace{0.5em}
\lordsday{21}

\catechismqa{54}{%
What do you believe concerning the holy catholic church?%
}{%
I believe that the Son of God, out of the whole human race, from the beginning of the world to its end, gathers, defends, and preserves for himself, by his Spirit and Word, in the unity of the true faith, a church chosen to everlasting life. And I believe that I am and forever shall remain a living member of it.%
}

\catechismqa{55}{%
What do you understand by the communion of saints?%
}{%
First, that believers, all and everyone, as members of Christ have communion with him and share in all his treasures and gifts. Second, that everyone is bound to use his gifts readily and cheerfully for the benefit and well-being of the other members.%
}

\catechismqa{56}{%
What do you believe concerning the forgiveness of sins?%
}{%
I believe that God, because of Christ's satisfaction, will no longer remember my sins, nor my sinful nature, against which I have to struggle all my life, but will graciously grant me the righteousness of Christ, that I may never come into condemnation.%
}
