%% gratitude-4.tex - Part 3d: Of Thankfulness (The Lord's Prayer)
%% Questions 120-129, Lord's Days XLVI-LII

%% ============================================================
%% LORD'S DAY XLVI
%% ============================================================

\lordsday{46}

\catechismqa{120}{%
Why has Christ commanded us to address God as Our Father?%
}{%
To awaken in us at the very beginning of our prayer that childlike reverence and trust toward God which should be the ground of our prayer: God has become our Father through Christ and will much less deny us what we ask of him in faith than our parents would refuse us earthly things.%
}

\catechismqa{121}{%
Why is there added, Who art in heaven?%
}{%
That we might have no earthly conception of the heavenly majesty of God, and that we might expect from his almighty power all things needed for body and soul.%
}

%% ============================================================
%% LORD'S DAY XLVII
%% ============================================================

\vspace{0.5em}
\lordsday{47}

\catechismqa{122}{%
What is the first petition?%
}{%
Hallowed be thy name. That is: Grant us first of all that we may rightly know thee, and sanctify, glorify, and praise thee in all thy works, in which shine forth thy almighty power, wisdom, goodness, righteousness, mercy, and truth. Grant us also that we may so direct our whole life---our thoughts, words, and actions---that thy name is not blasphemed because of us but always honoured and praised.%
}

%% ============================================================
%% LORD'S DAY XLVIII
%% ============================================================

\vspace{0.5em}
\lordsday{48}

\catechismqa{123}{%
What is the second petition?%
}{%
Thy kingdom come. That is: So rule us by thy Word and Spirit that more and more we submit to thee. Preserve and increase thy church. Destroy the works of the devil, every power that raises itself against thee, and every conspiracy against thy holy Word. Do all this until the fullness of thy kingdom comes, wherein thou shalt be all in all.%
}

%% ============================================================
%% LORD'S DAY XLIX
%% ============================================================

\vspace{0.5em}
\lordsday{49}

\catechismqa{124}{%
What is the third petition?%
}{%
Thy will be done, on earth as it is in heaven. That is: Grant that we and all men may deny our own will, and without any murmuring obey thy will, for it alone is good. Grant also that everyone may carry out the duties of his office and calling as willingly and faithfully as the angels in heaven.%
}

%% ============================================================
%% LORD'S DAY L
%% ============================================================

\vspace{0.5em}
\lordsday{50}

\catechismqa{125}{%
What is the fourth petition?%
}{%
Give us this day our daily bread. That is: Provide us with all our bodily needs so that we may acknowledge that thou art the only source of all that is good, and that without thy blessing neither our care and labour nor thy gifts can do us any good. Grant therefore that we may withdraw our trust from all creatures and place it in thee alone.%
}

%% ============================================================
%% LORD'S DAY LI
%% ============================================================

\vspace{0.5em}
\lordsday{51}

\catechismqa{126}{%
What is the fifth petition?%
}{%
Forgive us our debts, as we also have forgiven our debtors. That is: For the sake of Christ's blood, do not impute to us, wretched sinners, any of our transgressions, nor the evil which still clings to us, as we also find this evidence of thy grace in us that we are fully determined wholeheartedly to forgive our neighbour.%
}

%% ============================================================
%% LORD'S DAY LII
%% ============================================================

\vspace{0.5em}
\lordsday{52}

\catechismqa{127}{%
What is the sixth petition?%
}{%
And lead us not into temptation, but deliver us from the evil one. That is: Since we are so weak that we cannot stand on our own for one moment, and besides, our sworn enemies---the devil, the world, and our own flesh---do not cease to attack us, wilt thou preserve and strengthen us by the power of thy Holy Spirit so that we may stand firm against them and not be defeated in this spiritual war, until we finally obtain complete victory.%
}

\catechismqa{128}{%
How do you conclude your prayer?%
}{%
For thine is the kingdom and the power and the glory forever. That is: All this we ask of thee because, as our King, having power over all things, thou art both willing and able to give us all that is good, and because not we but thy holy name should receive all the glory forever.%
}

\catechismqa{129}{%
What does the word Amen mean?%
}{%
Amen means: It is true and certain. For God has much more certainly heard my prayer than I feel in my heart that I desire this of him.%
}
