%% gratitude-2.tex - Part 3b: Of Thankfulness (The Law and Prayer)
%% Questions 108-129, Lord's Days XLI-LII

%% ============================================================
%% LORD'S DAY XLI
%% ============================================================

\lordsday{41}

\catechismqa{108}{%
What does the seventh commandment teach us?%
}{%
That all unchastity is cursed by God. We must therefore detest it from the heart and live chaste and disciplined lives, both within and outside of holy marriage.%
}

\catechismqa{109}{%
Does God in this commandment forbid nothing more than adultery and similar shameful sins?%
}{%
Since we, body and soul, are temples of the Holy Spirit, it is God's will that we keep ourselves pure and holy. Therefore he forbids all unchaste acts, gestures, words, thoughts, desires, and whatever may entice us to unchastity.%
}

%% ============================================================
%% LORD'S DAY XLII
%% ============================================================

\vspace{0.5em}
\lordsday{42}

\catechismqa{110}{%
What does God forbid in the eighth commandment?%
}{%
God forbids not only theft and robbery which the authorities punish, but God also labels as theft all wicked tricks and schemes by which we seek to get for ourselves our neighbour's goods, whether by force or under the appearance of right, such as false weights and measures, deceptive advertising or merchandising, counterfeit money, or usury; also all greed and wasteful abuse of his gifts.%
}

\catechismqa{111}{%
What does God require of you in this commandment?%
}{%
That I promote my neighbour's good wherever I can and may, deal with him as I would like others to deal with me, and work faithfully so that I may be able to give to those in need.%
}

%% ============================================================
%% LORD'S DAY XLIII
%% ============================================================

\vspace{0.5em}
\lordsday{43}

\catechismqa{112}{%
What is required in the ninth commandment?%
}{%
That I bear false witness against no one, twist no one's words, be no backbiter or slanderer, join in condemning no one unheard or rashly. Rather, I must avoid, on pain of God's heavy wrath, all lying and deceit as the devil's own works. In court and everywhere else, I must love the truth, speak and confess it honestly, and do what I can to defend and promote my neighbour's honour and reputation.%
}

%% ============================================================
%% LORD'S DAY XLIV
%% ============================================================

\vspace{0.5em}
\lordsday{44}

\catechismqa{113}{%
What does the tenth commandment require of us?%
}{%
That not even the slightest thought or desire contrary to any of God's commandments should ever arise in our heart. Rather, we should always hate all sin with all our heart and delight in all righteousness.%
}

\catechismqa{114}{%
But can those converted to God keep these commandments perfectly?%
}{%
No. Even the holiest of them make only a small beginning in this obedience in this life. Nevertheless, they begin with earnest purpose to live not only according to some but according to all the commandments of God.%
}

\catechismqa{115}{%
If in this life no one can keep the ten commandments perfectly, why does God have them preached so strictly?%
}{%
First, that all our life long we may more and more come to know our sinful nature, and so become the more eager to seek the forgiveness of sins and righteousness in Christ. Second, that we may continually strive and ask God for the grace of the Holy Spirit, so that more and more we may be renewed in the image of God, until after this life we reach the goal of perfection.%
}

%% ============================================================
%% LORD'S DAY XLV
%% ============================================================

\vspace{0.5em}
\lordsday{45}

\catechismqa{116}{%
Why is prayer necessary for Christians?%
}{%
Because it is the chief part of the thankfulness which God requires of us, and because God will give his grace and Holy Spirit only to those who constantly and with heartfelt longing ask him for these gifts and thank him for them.%
}

\catechismqa{117}{%
What belongs to a prayer which pleases God and is heard by him?%
}{%
First, that we sincerely call upon the one true God only, who has revealed himself in his Word, for all that he has commanded us to pray. Second, that we thoroughly know our need and misery, so that we may humble ourselves before God. Third, that we rest on this firm foundation, that although we do not deserve it, God will certainly hear our prayer for the sake of Christ our Lord, as he has promised us in his Word.%
}

\catechismqa{118}{%
What has God commanded us to ask of him?%
}{%
All the things we need for body and soul, as included in the prayer which Christ our Lord himself taught us.%
}

\catechismqa{119}{%
What is the Lord's Prayer?%
}{%
Our Father who art in heaven, hallowed be thy name. Thy kingdom come. Thy will be done on earth as it is in heaven. Give us this day our daily bread. And forgive us our debts, as we also have forgiven our debtors. And lead us not into temptation, but deliver us from the evil one. For thine is the kingdom and the power and the glory forever. Amen.%
}

%% ============================================================
%% LORD'S DAY XLVI
%% ============================================================

\vspace{0.5em}
\lordsday{46}

\catechismqa{120}{%
Why has Christ commanded us to address God as Our Father?%
}{%
To awaken in us at the very beginning of our prayer that childlike reverence and trust toward God which should be the ground of our prayer: God has become our Father through Christ and will much less deny us what we ask of him in faith than our parents would refuse us earthly things.%
}

\catechismqa{121}{%
Why is there added, Who art in heaven?%
}{%
That we might have no earthly conception of the heavenly majesty of God, and that we might expect from his almighty power all things needed for body and soul.%
}

%% ============================================================
%% LORD'S DAY XLVII
%% ============================================================

\vspace{0.5em}
\lordsday{47}

\catechismqa{122}{%
What is the first petition?%
}{%
Hallowed be thy name. That is: Grant us first of all that we may rightly know thee, and sanctify, glorify, and praise thee in all thy works, in which shine forth thy almighty power, wisdom, goodness, righteousness, mercy, and truth. Grant us also that we may so direct our whole life---our thoughts, words, and actions---that thy name is not blasphemed because of us but always honoured and praised.%
}

%% ============================================================
%% LORD'S DAY XLVIII
%% ============================================================

\vspace{0.5em}
\lordsday{48}

\catechismqa{123}{%
What is the second petition?%
}{%
Thy kingdom come. That is: So rule us by thy Word and Spirit that more and more we submit to thee. Preserve and increase thy church. Destroy the works of the devil, every power that raises itself against thee, and every conspiracy against thy holy Word. Do all this until the fullness of thy kingdom comes, wherein thou shalt be all in all.%
}

%% ============================================================
%% LORD'S DAY XLIX
%% ============================================================

\vspace{0.5em}
\lordsday{49}

\catechismqa{124}{%
What is the third petition?%
}{%
Thy will be done, on earth as it is in heaven. That is: Grant that we and all men may deny our own will, and without any murmuring obey thy will, for it alone is good. Grant also that everyone may carry out the duties of his office and calling as willingly and faithfully as the angels in heaven.%
}

%% ============================================================
%% LORD'S DAY L
%% ============================================================

\vspace{0.5em}
\lordsday{50}

\catechismqa{125}{%
What is the fourth petition?%
}{%
Give us this day our daily bread. That is: Provide us with all our bodily needs so that we may acknowledge that thou art the only source of all that is good, and that without thy blessing neither our care and labour nor thy gifts can do us any good. Grant therefore that we may withdraw our trust from all creatures and place it in thee alone.%
}

%% ============================================================
%% LORD'S DAY LI
%% ============================================================

\vspace{0.5em}
\lordsday{51}

\catechismqa{126}{%
What is the fifth petition?%
}{%
Forgive us our debts, as we also have forgiven our debtors. That is: For the sake of Christ's blood, do not impute to us, wretched sinners, any of our transgressions, nor the evil which still clings to us, as we also find this evidence of thy grace in us that we are fully determined wholeheartedly to forgive our neighbour.%
}

%% ============================================================
%% LORD'S DAY LII
%% ============================================================

\vspace{0.5em}
\lordsday{52}

\catechismqa{127}{%
What is the sixth petition?%
}{%
And lead us not into temptation, but deliver us from the evil one. That is: Since we are so weak that we cannot stand on our own for one moment, and besides, our sworn enemies---the devil, the world, and our own flesh---do not cease to attack us, wilt thou preserve and strengthen us by the power of thy Holy Spirit so that we may stand firm against them and not be defeated in this spiritual war, until we finally obtain complete victory.%
}

\catechismqa{128}{%
How do you conclude your prayer?%
}{%
For thine is the kingdom and the power and the glory forever. That is: All this we ask of thee because, as our King, having power over all things, thou art both willing and able to give us all that is good, and because not we but thy holy name should receive all the glory forever.%
}

\catechismqa{129}{%
What does the word Amen mean?%
}{%
Amen means: It is true and certain. For God has much more certainly heard my prayer than I feel in my heart that I desire this of him.%
}
