%% guilt.tex - Part 1: Of Man's Misery
%% Questions 1-11, Lord's Days I-IV
%% Text from Philip Schaff's 1877 translation (public domain)

%% ============================================================
%% LORD'S DAY I
%% ============================================================

\lordsday{1}

\catechismqa{1}{%
What is thy only comfort in life and in death?%
}{%
That I, with body and soul, both in life and in death, am not my own,
but belong to my faithful Saviour Jesus Christ, who with His precious
blood has fully satisfied for all my sins, and redeemed me from all
the power of the devil; and so preserves me that without the will of
my heavenly Father not a hair can fall from my head; yea, that all
things must work together for my salvation. Wherefore, by His Holy
Spirit, He also assures me of eternal life, and makes me heartily
willing and ready henceforth to live unto Him.%
}

\catechismqa{2}{%
How many things are necessary for thee to know, that thou in this
comfort mayest live and die happily?%
}{%
Three things: First, the greatness of my sin and misery. Second, how
I am redeemed from all my sins and misery. Third, how I am to be
thankful to God for such redemption.%
}

%% ============================================================
%% LORD'S DAY II
%% ============================================================

\lordsday{2}

\catechismqa{3}{%
Whence knowest thou thy misery?%
}{%
Out of the Law of God.%
}

\catechismqa{4}{%
What doth the Law of God require of us?%
}{%
Christ teaches us that in sum, in Matthew 22:37--40: Thou shalt love
the Lord thy God with all thy heart, and with all thy soul, and with
all thy mind, and with all thy strength. This is the first and great
commandment; and the second is like unto it: Thou shalt love thy
neighbor as thyself. On these two commandments hang all the Law and
the Prophets.%
}

\catechismqa{5}{%
Canst thou keep all this perfectly?%
}{%
No; for I am by nature prone to hate God and my neighbor.%
}

%% ============================================================
%% LORD'S DAY III
%% ============================================================

\columnbreak
\lordsday{3}

\catechismqa{6}{%
Did God create man thus wicked and perverse?%
}{%
No; but God created man good, and after His own image, that is, in
righteousness and true holiness, that he might rightly know God his
Creator, heartily love Him, and live with Him in eternal blessedness,
to praise and glorify Him.%
}

\catechismqa{7}{%
Whence then comes this depraved nature of man?%
}{%
From the fall and disobedience of our first parents, Adam and Eve, in
Paradise, whereby our nature became so corrupt that we are all
conceived and born in sin.%
}

\catechismqa{8}{%
But are we so far depraved that we are wholly unapt to any good, and
prone to all evil?%
}{%
Yes; unless we are born again by the Spirit of God.%
}

%% ============================================================
%% LORD'S DAY IV
%% ============================================================

\lordsday{4}

\catechismqa{9}{%
Doth not God, then, wrong man by requiring of him in His Law that
which he cannot perform?%
}{%
No; for God so made man that he could perform it; but man, through
the instigation of the devil, by wilful disobedience deprived himself
and all his posterity of those divine gifts.%
}

\catechismqa{10}{%
Will God suffer such disobedience and apostasy to go unpunished?%
}{%
By no means; but He is terribly displeased with our inborn as well as
our actual sins, and will punish them in just judgment in time and
eternity, as He has declared: Cursed is every one that continueth not
in all things which are written in the book of the Law, to do them.%
}

\catechismqa{11}{%
But is not God also merciful?%
}{%
God is indeed merciful, but He is likewise just; wherefore His
justice requires that sin, which is committed against the most high
majesty of God, be also punished with extreme, that is, with
everlasting punishment both of body and soul.%
}
